---
title: "Stage | Analyse descriptive des données et modèle de survie "
author: "Catherineau Alexandre"
date: "`r Sys.Date()`"
output: html_document
---

# Code R stage

### Data and packages loading

```{r}
library(rlang)
library(dplyr)
data3C=read.table("base_Alexandre.txt",header=T)
head(data3C)
```

### Looking data and variables

```{r}
summary(data3C)
```

We select the center for each person from NUM.

```{r}
data3C$CENTRE[data3C$NUM>=10000]=1
data3C$CENTRE[data3C$NUM>=20000]=2
data3C$CENTRE[data3C$NUM>=30000]=3
```

We discretize the DIPNIV variable to make it easier to use.

```{r}
data3C$DIPNIVt[data3C$DIPNIV==1|data3C$DIPNIV==2|data3C$DIPNIV==3]=0
data3C$DIPNIVt[data3C$DIPNIV==4|data3C$DIPNIV==5]=1
```

We discretize the Apoe4 variable.

```{r}
data3C$score14t=cut(data3C$score14,c(0,6,8,10,14))
```

## Data manipulation to obtain the cross-sectional base

Package to obtain variables summary

```{r}
library(skimr)
data3C$DEM0=as.factor(data3C$DEM0)
data3C$DIPNIVt=as.factor(data3C$DIPNIVt)
data3C$DIPNIV=as.factor(data3C$DIPNIV)
data3C$DC=as.factor(data3C$DC)
data3C$SEXE=as.factor(data3C$SEXE)
data3C$avccorr0=as.factor(data3C$avccorr0)
data3C$malcorot=as.factor(data3C$malcorot)
data3C$APOE4=as.factor(data3C$APOE4)
data3C$CENTRE=as.factor(data3C$CENTRE)
```

The database *datatrans* is the transversal database at the first follow-up.

```{r}
excluSuivi=filter(data3C,SUIVI==8&abs(AGE0-AGEFIN)<0.0005)
excluD=filter(data3C,SUIVI==0&DEM0==1)
datatrans <- data3C %>% filter(SUIVI==0) %>% select(-DC,-DEM0,-SUIVI,-AGEFIN,-AGE,-AGEDEM)
```

We exclude all the people who don't have data for score14 variable. *#flow chart*

```{r}
data1 <- subset(datatrans, is.na(score14)==F)
```

We exclude all the people who don't have data for adjustment variables. *#flow chart*

```{r}
data2 <- subset(data1, is.na(DIPNIVt)==F & is.na(APOE4)==F & is.na(avccorr0)==F & is.na(malcorot)==F)
```

**Tables creation with incidents events.** (death, dementia, end)

-   Final age

```{r}
longage = data3C %>% group_by(ID) %>% arrange(desc(AGEFIN)) %>% filter(row_number()==1) %>% select(ID,AGEFIN)
```

-   Death

```{r}
longdc = data3C %>% group_by(ID) %>% arrange(desc(AGEFIN)) %>% filter(row_number()==1) 

#Here people not deceased -> age end censored
longdc0 = longdc %>% group_by(ID) %>% filter(DC==0) %>% select(ID,DC,AGEFIN) %>% mutate(AGEFIN=max(AGEFIN))

#Here people deceased -> age of death
longdc1 = longdc %>% group_by(ID) %>% filter(DC==1) %>% select(ID,DC,AGEFIN)

longdc = union(longdc1,longdc0)
longdc = rename(longdc, AGEDC = AGEFIN)
```

-   Dementia

```{r}
#Here people without dementia -> last agedem as dementia age
longdem0 = data3C %>% group_by(ID) %>% arrange(desc(AGEDEM))  %>% filter(row_number()==1) %>% filter(DEM0==0) %>% select(ID,DEM0,AGEDEM)

#Here people with dementia -> dementia age
longdem1=data3C %>% group_by(ID) %>% filter(DEM0==1) %>% filter(row_number()==1)  %>% select(ID,DEM0,AGEDEM)

longdem=union(longdem1,longdem0)
```

Now we **merge** the three tables with ID.

```{r}
datatransvers=data2 %>% left_join(longdem,by="ID") %>% left_join(longdc,by="ID") %>% left_join(longage,by="ID")
```

Removal of lines where people have dementia or a history of cardiovascular disease at follow-up 0. *#flow chart*

```{r}
excluSuivi=(datatransvers %>% select(ID)) %>% intersect((excluSuivi %>% select(ID)))

excludedDEM=subset(datatransvers, abs(AGE0-AGEDEM)<0.0005)

excludedcardio=subset(datatransvers, avccorr0==1 | malcorot==1|malcorot==2)

datatransvers <- subset(datatransvers, abs(AGE0-AGEDEM)>0.0005) 

datatransvers <- subset(datatransvers,avccorr0==0 & malcorot==0)
```

Final data ---\> *datatransvers*

**Flow chart**

```{r}
library(Gmisc)
library(magrittr)
library(glue)
 
# The key boxes that we want to plot
Pop <- boxGrob(glue("Population globale",
            "n = {pop}",
            pop = txtInt(length(datatrans$ID)),
            .sep = "\n"))
databox1 <- boxGrob(glue("Echantillon avec\nun score cardiovasculaire",
            "n = {pop}",
            pop = txtInt(length(data1$ID)),
            .sep = "\n"))
databox2 <- boxGrob(glue("Echantillon avec les\nvariables d'ajustement",
            "n = {pop}",
            pop = txtInt(length(data2$ID)),
            .sep = "\n"))
databox3 <- boxGrob(glue("Base finale",
            "n = {pop}",
            pop = txtInt(length(datatransvers$ID)),
            .sep = "\n"))
excluded1 <- boxGrob(glue("Scores\ncardiovasculaires\n manquants (n = {tot})",
          tot=txtInt(length(datatrans$ID)-length(data1$ID)),
            .sep = "\n",
            just = "left"))
excluded2 <- boxGrob(glue("Variables d'ajustement\nmanquantes (n = {tot}) :",
            "- Niveau d'étude (n={DIPNIV})",
            "- APOE4 (n={APOE4})",
            "- Antécédents cardiovasculaires (n={ant})",
            DIPNIV = sum(is.na(data1$DIPNIVt)),
            APOE4 = sum(is.na(data1$APOE4)),
            ant = length(data1$ID)-length(data2$ID)-DIPNIV-APOE4,
            tot=txtInt(length(data1$ID)-length(data2$ID)),
            .sep = "\n",
            just = "left"))
excluded3 <- boxGrob(glue("Au suivi 0, personnes :",
            "- démentes (n={DEM})",
            "- présentant des antécédents\ncardiovasculaires (n={avc_mal})",
            DEM = length(excludedDEM$ID),
            avc_mal = length(data2$ID)-length(datatransvers$ID)-length(excludedDEM$ID),
            .sep = "\n",
            just = "left"))
 
# Move boxes to where we want them
vert <- spreadVertical(Pop=Pop,
                       databox1 = databox1,
                       databox2 = databox2,
                       databox3 = databox3)
 
y1 <- coords(vert$databox1)$top +
  distance(vert$Pop, vert$databox1, half = TRUE, center = FALSE)
y2 <- coords(vert$databox2)$top +
  distance(vert$databox1, vert$databox2, half = TRUE, center = FALSE)
y3 <- coords(vert$databox3)$top +
  distance(vert$databox2, vert$databox3, half = TRUE, center = FALSE)

excluded1 <- moveBox(excluded1,
                    x = .8,
                    y = y1)
excluded2 <- moveBox(excluded2,
                    x = .8,
                    y = y2)
excluded3 <- moveBox(excluded3,
                    x = .8,
                    y = y3)

# Connect vertical arrows, skip last box
for (i in 1:(length(vert) - 1)) {
  connectGrob(vert[[i]], vert[[i + 1]], type = "vert") %>%
    print
}
 
# Add a connection to the exclusions
connectGrob(vert$databox1, excluded2, type = "L")
connectGrob(vert$Pop, excluded1, type = "L")
connectGrob(vert$databox2, excluded3, type = "L")

# Print boxes
vert
excluded1
excluded2
excluded3
```

Calculation of low and upper for the 4 cases.

We take data3C but only with people already included.

```{r}
df=semi_join(data3C,datatransvers, by = "ID")
```

Demented

```{r}
#Lower age
low = df %>% group_by(ID) %>% arrange(desc(AGE)) %>% filter(DEM0==0) %>% filter(row_number()==1) %>% select(AGE,ID) %>% rename(AGEDEMlow=AGE)

#Upper age
upper=df %>% group_by(ID)%>% arrange(AGE) %>% filter(DEM0==1) %>% filter(row_number()==1) %>%select(AGE,ID) %>% rename(AGEDEMup=AGE)

D=inner_join(low,upper,by="ID")
```

Non demented non deceased

```{r}
#Lower age 
low = datatransvers %>% filter(DEM0==0&DC==0) %>% select(ID,AGEDEM) %>% rename(AGEDEMlow=AGEDEM)

#Upper age 
upper = datatransvers %>% filter(DEM0==0&DC==0) %>% select(ID,AGEDEM) %>% rename(AGEDEMup=AGEDEM)

nDnD=inner_join(low,upper,by="ID")
```

non Déments décédés

```{r}
#Lower age
low = datatransvers %>% filter(DEM0==0&DC==1) %>% select(ID,AGEDEM) %>% rename(AGEDEMlow=AGEDEM)

#Upper age
upper = datatransvers %>% filter(DEM0==0&DC==1) %>% select(ID,AGEDC) %>% rename(AGEDEMup=AGEDC)

nDD=inner_join(low,upper,by="ID")
```

```{r}
df=union(D,union(nDD,nDnD))
datatransvers= datatransvers %>% left_join(df,by="ID")
```

```{r}
df=data3C %>% group_by(ID) %>% semi_join(excludedDEM,by="ID")
```

Here this patient has his dementia age increasing after being diagnosed as demented. We assume that his age was not meant to be increased.

```{r}
datatransvers$AGEDEMup[3251]=(data3C%>%filter(ID==4954&SUIVI==5))$AGEDEM
```

## Survival methods

Delays calculation

```{r}
datatransvers$DELAIDEM=datatransvers$AGEDEM-datatransvers$AGE0
datatransvers$DELAIDC=datatransvers$AGEDC-datatransvers$AGE0
```

Database writing

```{r}
write.table(datatransvers,file="datatransvers.txt", fileEncoding = "UTF-8")
```
