---
title: "Intervalle_conf"
author: "Catherineau Alexandre"
date: "`r Sys.Date()`"
output: html_document
---

### Data and packages loading

```{r}
data=read.table("datatransvers.txt",header=T)
library(dplyr)
library(tidyr)
library(survival)
library(pbapply)
library(ggplot2)
```

### Variable arrangement 

#### Reduced centered age

```{r}
data$AGE0t=(data$AGE0-mean(data$AGE0))/sd(data$AGE0)
```

#### Stratification apoe4 x Centre

```{r}
data$Apoecentre[data$APOE4==0&data$CENTRE==1]=1
data$Apoecentre[data$APOE4==0&data$CENTRE==2]=2
data$Apoecentre[data$APOE4==0&data$CENTRE==3]=3
data$Apoecentre[data$APOE4==1&data$CENTRE==1]=4
data$Apoecentre[data$APOE4==1&data$CENTRE==2]=5
data$Apoecentre[data$APOE4==1&data$CENTRE==3]=6
```

### Selection of persons

```{r}
selection <- function(n=1000){
  echantillon=data[sample(1:nrow(data),n,replace=TRUE),]
  return(echantillon)
}
d=selection()
```

### Calculation cumulative risk Mdem

```{r}
Rcumuldem <- function(M,score,sexe=1,dipniv=0,apoe4=0,age0=0,CENTRE=1){
  data0=data.frame("SEXE"=sexe,"DIPNIVt"=dipniv,"AGE0t"=age0,"score14"=score,"Apoecentre"=CENTRE)
  hazM=basehaz(M,newdata=data0)
  return(hazM)
}
```

### Calculation cumulative risk Mdc & Mdcdem

```{r}
Rcumul <- function(M,score,sexe=1,dipniv=0,apoe4=0,age0=0,CENTRE=1){
  data0=data.frame("SEXE"=sexe,"DIPNIVt"=dipniv,"AGE0t"=age0,"score14"=score,"APOE4"=apoe4,"CENTRE"=CENTRE)
  hazM=basehaz(M,newdata=data0)
  return(hazM)
}
```

### Probability of being healthy

```{r}
Fsain <- function(t,L){

  hazdc <- L %>% arrange(abs(time-t)) %>% filter(row_number()==1) %>% pull(hazdc)
  
  hazdem<- L %>% arrange(abs(time-t)) %>% filter(row_number()==1) %>% pull(hazdem)
  
  proba_sain = exp(-hazdem-hazdc)
  return(proba_sain)
}
```

### Instantaneous risk

```{r}
Rinstant <- function(Mdem,Mdc,Mdcdem,score){
  L=Rcumuldem(Mdem,score)%>% select(time,hazard)
  Lambda0=L
  
  time=diff(L$time)
  hazard=diff(L$hazard)
  hazard <- c(hazard, tail(hazard, 1))
  time <- c(time, tail(time, 1))
  
  Lambda0$hazard=hazard/time
  
  return(Lambda0)
}
```

### Data frame with all times

```{r}
Ftable <- function(Mdem,Mdc,Mdcdem,df,a,abis){
#calculation of hazdem
Ldem2=Rcumuldem(Mdem,abis)
Ldem2=Ldem2 %>% rename(hazdem=hazard) %>% select(time,hazdem)

#calculation of hazdc
Ldc2=Rcumul(Mdc,a)
Ldc2=Ldc2 %>% rename(hazdc=hazard) %>% select(time,hazdc)

#calculation of hazdemdc
Ldcdem2=Rcumul(Mdcdem,a)
Ldcdem2=Ldcdem2 %>% rename(hazdcdem=hazard) %>% select(time,hazdcdem)

#base creation with all the times
#tdem, tdc, tdemdc
L = (Ldem2 %>% select(time)) %>% bind_rows(Ldc2 %>% select(time)) %>% bind_rows(Ldcdem2 %>% select(time)) %>% unique()

#Filling of hazdem
Ldem = Ldem2 %>% full_join(L,by='time',relationship= "many-to-many") %>% arrange(time) %>% fill(hazdem,.direction='downup') %>% unique()

#Filling of hazdc
Ldc = Ldc2 %>% full_join(L,by='time',relationship= "many-to-many") %>% arrange(time) %>% fill(hazdc,.direction='downup')

#Filling of hazdemdc(a)
Ldcdem = Ldcdem2 %>% full_join(L,by='time',relationship= "many-to-many") %>% arrange(time) %>% fill(hazdcdem,.direction='downup')

#Instantaneous risk calculation alphadem
Lalphadem=Rinstant(Mdem,Mdc,Mdcdem,abis) 
Lalphadem=Lalphadem %>% rename(alphadem=hazard)

#Filling of alphadem
Lalphadem = Lalphadem %>% full_join(L,by='time',relationship= "many-to-many") %>% arrange(time) %>% fill(alphadem,.direction='downup')

#Regrouping of hazdem, hazdc, hazdemdc and alphadem
L = (((Ldem %>% inner_join(Ldc,by='time',relationship= "many-to-many")) %>% inner_join(Ldcdem,by='time',relationship= "many-to-many")) %>%  inner_join(Lalphadem,by='time',relationship= "many-to-many"))

return(L)
}
```

### Survival function

```{r}
Fsurvie <- function(Mdem,Mdc,Mdcdem,L,t=10,a=0,abis=0,sexe=1,dipniv=0,apoe4=0,age0=0){
  i=1
  calcul=0
  #L=Ftable(Mdem,Mdc,Mdcdem,df,a,abis)
  #We take the closest time
  hazdcdem2 <- L %>% arrange(abs(time-t)) %>% filter(row_number()==1) %>% pull(hazdcdem)
  
  #Calculation
  while(L$time[i]<t){
    calcul = calcul+(exp(-L$hazdem[i]-L$hazdc[i])*L$alphadem[i]*exp(-hazdcdem2+L$hazdcdem[i]))*(min(t,L$time[i+1])-L$time[i])
    i=i+1
  }

  proba_sain=Fsain(t,L)
  #return(calcul)
  return(calcul+proba_sain)
}
```

## Bootstrap survival

```{r}
Bootstrap_surv <- function(B=100,n=6000,a=10,abis=10,tb=0){
  est_survie <- data.frame(matrix(nrow = 17, ncol = B))
  #est_survie=data.frame()
  for(j in 1:B){
    if(tb==0){df=selection(n)}
    else{df=data}
    
    #Dementia model
    Mdem=coxph(Surv(DELAIDEM,DEM0==1)~score14+SEXE+DIPNIVt+strata(Apoecentre)+AGE0t,data=df)
    
    #Death model without dementia
    dfdem = df %>% mutate(AGEDCt=if_else(DEM0==0,AGEDC,AGEDEM),DELAIDCt=AGEDCt-AGE0,DCt=if_else(DEM0==0,DC,0))
    Mdc=coxph(Surv(DELAIDCt,DCt==1)~score14+SEXE+DIPNIVt+APOE4+AGE0t+strata(CENTRE),data=dfdem)
    
    #Death model with dementia
    Mdcdem=coxph(Surv(DELAIDEM,DELAIDC,DC==1)~score14+SEXE+DIPNIVt+APOE4+AGE0t+strata(CENTRE),data=df%>% filter(DEM0==1))
    L=Ftable(Mdem,Mdc,Mdcdem,df,a,abis)
    # for(t in 1:17){
    #    est_survie[j,t]=Fsurvie(Mdem,Mdc,Mdcdem,L,t,a,abis)
    #  }
    est_survie[j] <- sapply(1:17, function(t){
      Fsurvie(Mdem, Mdc, Mdcdem, L, t, a, abis)
    })
    print(j)
  }
  return(est_survie)
}
survie=Bootstrap_surv(5,6197,12,12)
```

## Bootstrap effect

```{r}
Bootstrap_effet <- function(B=10,n=6000,a=12,abis=4,tb=0){
  est_part=data.frame()
  for(j in 1:B){
    if(tb==0){df=selection(n)}
    else{df=data}
    
    #Dementia model
    Mdem=coxph(Surv(DELAIDEM,DEM0==1)~score14+SEXE+DIPNIVt+strata(Apoecentre)+AGE0t,data=df)
    
    #Death model without dementia
    dfdem = df %>% mutate(AGEDCt=if_else(DEM0==0,AGEDC,AGEDEM),DELAIDCt=AGEDCt-AGE0,DCt=if_else(DEM0==0,DC,0))
    Mdc=coxph(Surv(DELAIDCt,DCt==1)~score14+SEXE+DIPNIVt+APOE4+AGE0t+strata(CENTRE),data=dfdem)
    
    #Detah model with dementia
    Mdcdem=coxph(Surv(DELAIDEM,DELAIDC,DC==1)~score14+SEXE+DIPNIVt+APOE4+AGE0t+strata(CENTRE),data=df%>% filter(DEM0==1))
    
    L1=Ftable(Mdem,Mdc,Mdcdem,df,a,a)
    L2=Ftable(Mdem,Mdc,Mdcdem,df,abis,abis)
    L3=Ftable(Mdem,Mdc,Mdcdem,df,abis,a)
    for(t in 1:17){
      est_total=Fsurvie(Mdem,Mdc,Mdcdem,L1,t,a,a)-Fsurvie(Mdem,Mdc,Mdcdem,L2,t,abis,abis)
      est_ind=Fsurvie(Mdem,Mdc,Mdcdem,L3,t,abis,a)-Fsurvie(Mdem,Mdc,Mdcdem,L2,t,abis,abis)
      est_part[j,t]=(est_ind/est_total)*100
    }
  print(j)
  }
  return(est_part)
}
effet=Bootstrap_effet(10,6197,12,4)
#quantile(survie,probs=c(0.025,0.975))
```

## Graph

### Confidence interval calculation, survival probability

```{r}
survie=Bootstrap_surv(2,6197,12,12)
survie_quartiles_sup <- apply(survie, 1, quantile, probs = c(0.025, 0.975))

survie=Bootstrap_surv(2,6197,8,8)
survie_quartiles_mid <- apply(survie, 1, quantile, probs = c(0.025, 0.975))

survie=Bootstrap_surv(2,6197,4,4)
survie_quartiles_inf <- apply(survie, 1, quantile, probs = c(0.025, 0.975))
```

### Confidence interval calculation, indirect effect

```{r}
effet=Bootstrap_effet(10,6197,12,4)
effet_quartiles <- effet %>% summarise_all(quantile, probs = c(0.025, 0.975))
```

### Graph survival probability

```{r}
#Calculation with the real dataset
Fi=data.frame()
Fi=data.frame(seq(1:17))
Fi$time=seq(1:17)
cvibas=Bootstrap_surv(2,NA,4,4,1)%>%rename(cvibas=X1)
Fi$cvibas=cvibas$cvibas
cvimilieu=Bootstrap_surv(2,NA,8,8,1)%>%rename(cvimilieu=X1)
Fi$cvimilieu=cvimilieu$cvimilieu
cvihaut=Bootstrap_surv(2,NA,12,12,1)%>%rename(cvihaut=X1)
Fi$cvihaut=cvihaut$cvihaut

#We take each CI
Fi$cviinfS=survie_quartiles_sup[1,]
Fi$cvisupS=survie_quartiles_sup[2,]
Fi$cviinfM=survie_quartiles_mid[1,]
Fi$cvisupM=survie_quartiles_mid[2,]
Fi$cviinfI=survie_quartiles_inf[1,]
Fi$cvisupI=survie_quartiles_inf[2,]
p <- ggplot(data = Fi, aes(x = time)) +
  geom_line(aes(y = cvibas, color = "Low score")) +
  geom_line(aes(y = cvimilieu, color = "Moderate score")) +
  geom_line(aes(y = cvihaut, color = "High score")) +
  scale_color_manual(values = rev(c("red", "orange", "green"))) +
  geom_ribbon(data = Fi, aes(x = time, ymin = cviinfS, ymax = cvisupS, fill = "High score"), alpha = 0.3) +
  geom_ribbon(data = Fi, aes(x = time, ymin = cviinfM, ymax = cvisupM, fill = "Moderate score"), alpha = 0.3) +
  geom_ribbon(data = Fi, aes(x = time, ymin = cviinfI, ymax = cvisupI, fill = "Low score"), alpha = 0.3) +
  ylab("Probability to be alive") +
  xlab("Time (years)") +
  theme_bw() +
  theme()+
  ylim(0, 1) 

p + labs(color = "Scores", fill = "95% confidence interval")+ scale_fill_manual(values = c("Low score" = "lightcoral", "Moderate score" = "lightsalmon", "High score" = "lightgreen"))
```

### Graph indirect effect proportion

```{r}
Fi=data.frame(seq(1:17))
Fi$time=seq(1:17)
part=t(Bootstrap_effet(2,NA,12,4,1))
colnames(part)=c("part","X2")
Fi$part=part[,1]

Fi$par_inf=t(effet_quartiles[1,])
Fi$par_sup=t(effet_quartiles[2,])

p <- ggplot(data = Fi, aes(x = time)) +
  geom_line(aes(y = part)) +
  geom_ribbon(aes(ymin = par_inf, ymax = par_sup, fill = "95% confidence interval"), alpha = 0.5) +
  ylab("Proportion Indirect Effect (%)") +
  xlab("Time (years)") +
  theme_bw() +
  ylim(-20, 110) +
  xlim(3, 17) +
  scale_fill_manual(values = "gray60", guide = guide_legend(title = "Legend"))
p
```
