---
title: "Rendu modèles de survie"
author: "Catherineau Alexandre"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  html_document theme:paper: default
  pdf_document: default
---

### Data loading

```{r}
data=read.table("datatransvers.txt",header=T)
```

### Loading the necessary packages

```{r, message=FALSE, warning = FALSE}
library(survminer, quietly = TRUE)
library(survival)
library(dplyr)
library(cowplot)
library(tidyr)
```

## Variable arrangement

### Reduced centered age

```{r}
data$AGE0t=(data$AGE0-mean(data$AGE0))/sd(data$AGE0)
```

### Stratification apoe4 x Centre

```{r}
data$Apoecentre[data$APOE4==0&data$CENTRE==1]=1
data$Apoecentre[data$APOE4==0&data$CENTRE==2]=2
data$Apoecentre[data$APOE4==0&data$CENTRE==3]=3
data$Apoecentre[data$APOE4==1&data$CENTRE==1]=4
data$Apoecentre[data$APOE4==1&data$CENTRE==2]=5
data$Apoecentre[data$APOE4==1&data$CENTRE==3]=6
```

------------------------------------------------------------------------

## Calculation Cumulative risk

Function :

```{r}
Rcumuldem <- function(M,score,sexe=1,dipniv=0,apoe4=0,age0=0,CENTRE=1){
  #We take lambda0 
  data0=data.frame("SEXE"=sexe,"DIPNIVt"=dipniv,"AGE0t"=age0,"score14"=score,"Apoecentre"=CENTRE)
  
  hazM=basehaz(M,newdata=data0)
  
  return(hazM)
}
```

```{r}
Rcumul <- function(M,score,sexe=1,dipniv=0,apoe4=0,age0=0,CENTRE=1){
  #We take lambda0 
  data0=data.frame("SEXE"=sexe,"DIPNIVt"=dipniv,"AGE0t"=age0,"score14"=score,"APOE4"=apoe4,"CENTRE"=CENTRE)
  hazM=basehaz(M,newdata=data0)
  
  return(hazM)
}
```

To calculate Lambda, we need to use this formula :

$$
𝚲^e(t)=𝛬_o^e(t)*exp( VAR_i*𝛽_i )
$$

------------------------------------------------------------------------

## Dementia model *(Mdem)*

### Model writing and rendering

```{r}
Mdem=coxph(Surv(DELAIDEM,DEM0==1)~score14+SEXE+DIPNIVt+strata(Apoecentre)+AGE0t,data=data)
Mdem
```

### Model graph

```{r}
ggsurvplot(survfit(Mdem),data=data,risk.table=TRUE,cumevents=TRUE,
          xlab="Time in years",
          ylab="Probability to become demented",fun="event",
          conf.int=T,
          ylim=c(0,1),
          xlim=c(0,16),
          break.time.by=1,
          censor=F)
a=survfit(Mdem)
```

------------------------------------------------------------------------

## Death model without dementia*(Mdc)*

### Model preparation

```{r}
data = data %>% mutate(AGEDCt=if_else(DEM0==0,AGEDC,AGEDEM),DELAIDCt=AGEDCt-AGE0,DCt=if_else(DEM0==0,DC,0))
```

### Model writing and rendering

```{r}
Mdc=coxph(Surv(DELAIDCt,DCt==1)~score14+SEXE+DIPNIVt+APOE4+AGE0t+strata(CENTRE),data=data)
Mdc
```

### Model graph

```{r}
ggsurvplot(survfit(Mdc),data=data,risk.table=TRUE,cumevents=TRUE,
          xlab="Time in years",
          ylab="Probability to become demented",fun="event",
          conf.int=T,
          ylim=c(0,1),
          xlim=c(0,16),
          break.time.by=1,
          censor=F)
```

------------------------------------------------------------------------

## Death model with dementia *(Mdcdemt) -\> semi-Markov*

--\> not used during the internship

### Model writing and rendering

```{r}
Mdcdemt=coxph(Surv(AGEDC-AGEDEM,DC==1)~score14+SEXE+DIPNIVt+APOE4+AGE0+strata(CENTRE),data=data %>% filter(DEM0==1))
Mdcdemt
```

###Model graph

```{r}
ggsurvplot(survfit(Mdcdemt),data=data,risk.table=TRUE,cumevents=TRUE,
          xlab="Time in years",
          ylab="Probability to become demented",fun="event",
          conf.int=T,
          ylim=c(0,1),
          xlim=c(0,16),
          break.time.by=1,
          censor=F)
```

## Death model with dementia *(Mdcdem)* -\> Markov

### Model writing and rendering

```{r}
Mdcdem=coxph(Surv(DELAIDEM,DELAIDC,DC==1)~score14+SEXE+DIPNIVt+APOE4+AGE0t+strata(CENTRE),data=data%>% filter(DEM0==1))
Mdcdem
```

### Model graph

```{r}
ggsurvplot(survfit(Mdcdem),data=data,risk.table=TRUE,cumevents=TRUE,
          xlab="Time in years",
          ylab="Probability to become demented",fun="event",
          conf.int=T,
          ylim=c(0,1),
          xlim=c(0,16),
          break.time.by=1,
          censor=F)
```

------------------------------------------------------------------------

## Probabilities

### Function probability of being healthy

```{r}
Fsain <- function(t,L){
  #We take the closest time
  hazdc <- L %>% arrange(abs(time-t)) %>% filter(row_number()==1) %>% pull(hazdc)
  
  hazdem<- L %>% arrange(abs(time-t)) %>% filter(row_number()==1) %>% pull(hazdem)
  
  proba_sain = exp(-hazdem-hazdc)
  return(proba_sain)
}
```

### Function instantaneous risk

```{r}
Rinstant <- function(score){
  #function instantaneous risk (alphaDEM)
  L=Rcumuldem(Mdem,score)
  L=L %>% select(time,hazard)
  Lambda0=L
  
  #calculation
  for(i in 1:length(L$hazard)-1){
    Lambda0$hazard[i]=(L$hazard[i+1]-L$hazard[i])/(L$time[i+1]-L$time[i])
  }
  
  return(Lambda0)
}
```

### Table creation

```{r}
Ftable <- function(a,abis){
#calculation of hazdem
Ldem2=Rcumuldem(Mdem,abis)
Ldem2=Ldem2 %>% rename(hazdem=hazard) %>% select(time,hazdem)

#calculation of hazdc
Ldc2=Rcumul(Mdc,a)
Ldc2=Ldc2 %>% rename(hazdc=hazard) %>% select(time,hazdc)

#calculation of hazdemdc
Ldcdem2=Rcumul(Mdcdem,a)
Ldcdem2=Ldcdem2 %>% rename(hazdcdem=hazard) %>% select(time,hazdcdem)

#Base creation with all times
#tdem, tdc, tdemdc
L = (Ldem2 %>% select(time)) %>% bind_rows(Ldc2 %>% select(time)) %>% bind_rows(Ldcdem2 %>% select(time)) %>% unique()

#Filling of hazdem
Ldem = Ldem2 %>% full_join(L,by='time',relationship= "many-to-many") %>% arrange(time) %>% fill(hazdem,.direction='downup') %>% unique()

#Filling of hazdc
Ldc = Ldc2 %>% full_join(L,by='time',relationship= "many-to-many") %>% arrange(time) %>% fill(hazdc,.direction='downup')

#Filling of hazdemdc(a)
Ldcdem = Ldcdem2 %>% full_join(L,by='time',relationship= "many-to-many") %>% arrange(time) %>% fill(hazdcdem,.direction='downup')

#Calcul du risque instantané alphadem
Lalphadem=Rinstant(abis) 
Lalphadem=Lalphadem %>% rename(alphadem=hazard)

#Filling of alphadem
Lalphadem = Lalphadem %>% full_join(L,by='time',relationship= "many-to-many") %>% arrange(time) %>% fill(alphadem,.direction='downup')

#Regrouping of hazdem, hazdc, hazdemdc and alphadem
L = (((Ldem %>% inner_join(Ldc,by='time',relationship= "many-to-many")) %>% inner_join(Ldcdem,by='time',relationship= "many-to-many")) %>%  inner_join(Lalphadem,by='time',relationship= "many-to-many"))

return(L)
}
```

### Survival function

```{r}
Fsurvie <- function(resultat=0,t=1,a=0,abis=0,sexe=1,dipniv=0,apoe4=0,age0=0){
  L=Ftable(a,abis)
  i=1
  calcul=0
  #We take the closest time
  hazdcdem2 <- L %>% arrange(abs(time-t)) %>% filter(row_number()==1) %>% pull(hazdcdem)
  
  while(L$time[i]<t){
    calcul = calcul+(exp(-L$hazdem[i]-L$hazdc[i])*L$alphadem[i]*exp(-hazdcdem2+L$hazdcdem[i]))*(min(t,L$time[i+1])-L$time[i])
    i=i+1
  }
  proba_sain=Fsain(t,L)
  if(resultat==0){
    return(calcul+proba_sain)
  }
  if(resultat==1){
    return(proba_sain)
  }
  if(resultat==2){
    return(calcul)
  }
}
Fsurvie(2,t=17,4,4)
```

### Graph survival probability

```{r}
Fi=data.frame(seq(1:17))
Fi$time=seq(1:17)
for(i in(1:17)){
  Fi$cvibas[i]=Fsurvie(0,i,4,4)
  Fi$cvimilieu[i]=Fsurvie(0,i,8,8)
  Fi$cvihaut[i]=Fsurvie(0,i,12,12)
}

p=ggplot(data = Fi, aes(x=time))+
  geom_line(aes(y=cvibas,color = "Low score"))+
  geom_line(aes(y=cvimilieu,color = "Moderate score"))+
  geom_line(aes(y=cvihaut,color = "High score"))+
  scale_color_manual(values = c("green", "orange","red"))
p + ylab("Survival probability") + xlab("Time (years)")+ theme_bw()+ylim(0,1)+ theme(legend.title = element_blank())
```

```{r}
Fi=data.frame(seq(1:17))
Fi$time=seq(1:17)
for(i in(1:17)){
  Fi$Sbas[i]=Fsurvie(1,i,4,4)
  Fi$Smilieu[i]=Fsurvie(1,i,8,8)
  Fi$Shaut[i]=Fsurvie(1,i,12,12)
}
for(i in(1:17)){
  Fi$Dbas[i]=Fsurvie(2,i,4,4)
  Fi$Dmilieu[i]=Fsurvie(2,i,8,8)
  Fi$Dhaut[i]=Fsurvie(2,i,12,12)
}

p=ggplot(data = Fi, aes(x=time))+
  geom_line(aes(y=Sbas,color = "Low score"))+
  geom_line(aes(y=Smilieu,color = "Moderate score"))+
  geom_line(aes(y=Shaut,color = "High score"))+
  scale_color_manual(values = c("green", "red","orange"))
p2=ggplot(data = Fi, aes(x=time))+
  geom_line(aes(y=Dbas,color = "Low score"))+
  geom_line(aes(y=Dmilieu,color = "Moderate score"))+
  geom_line(aes(y=Dhaut,color = "High score"))+
  scale_color_manual(values = c("green", "orange","red"))
p + ylab("Probability to be alive without dementia") + xlab("Time (years)")+ theme_bw()+ylim(0,1)+ theme(legend.title = element_blank())

p2 + ylab("Probabilité d'être vivant et dément") + xlab("Temps (années)")+ theme_bw()+ylim(0,1)+ theme(legend.title = element_blank())
```

### Calculating direct and indirect effects

```{r}
Fi=data.frame(seq(1:17))
Fi$time=seq(1:17)
for(i in(1:17)){
  Fi$effetotal[i]=Fsurvie(0,i,12,12)-Fsurvie(0,i,4,4)
  Fi$effetindirect[i]=Fsurvie(0,i,4,12)-Fsurvie(0,i,4,4)
  #a=4 et abis=12 puis a=4 et abis=4
  Fi$part[i]=(Fi$effetindirect[i]/Fi$effetotal[i])*100
}
#library(openxlsx)
#write.xlsx(x = Fi, file = "data_ex_export.xlsx")
```

```{r}
p=ggplot(data = Fi, aes(x=time))+
  geom_line(aes(y=part))
p + ylab("Proportion Indirect Effect (%)") + xlab("Time (years)")+ theme_bw()+ylim(0,100)+xlim(3,17) 
```

## Model adequacy

### Proportionality of risks

-\> Observations must be independent, and the risk rate must be constant over time. This means that the proportionality of risk from one observation to another must not vary over time.

```{r}
prop=cox.zph(Mdcdem,transform='identity')
print(prop)
```

Here problems with Mdc and Mdem Mdem because of apoe4 and Mdc because of AGE0

Hypothesis not validated as there is significant interaction with time

global p_value less than 0.05 for Mdc and Mdem

```{r}
MdemT=coxph(Surv(DELAIDEM,DEM0==1)~score14+SEXE+DIPNIVt+strata(APOE4)+AGE0,data=data)
MdemT
prop=cox.zph(MdemT,transform='identity')
print(prop)
```

By stratifying with apoe4 it seems that the problem has been solved.

### Log-linearity

The relative risk, for a one-unit increase in the variable, remains the same whatever the value of the explanatory variable.

Spline model

```{r}
Mdem2=coxph(Surv(DELAIDCt,DCt==1)~pspline(score14)+AGE0+SEXE+DIPNIVt+APOE4,data=data)
Mdem2
```
